Debugging by \textit{lastChange}

### 2010-11-18 papers/pldi2011/sigplanconf-template.tex

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{algorithm}

@@

By browsing through the code, the developer can quickly determine that the value displayed on the 
button is set at line 22. To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, execution is paused at line
22. Figure ~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

@@

In contrast, we have added a high-level function in debugger,
\textit{lastChange}, which provides the answer without tedious manual
effort from the developer. By right clicking on
\texttt{myObject.myProperty} in the Watch panel, the developer can run
\textit{lastChange} command (Figure ~\ref{fig:example2}). Debugger
re-executes the program and halts again at the breakpoint on line 22.
However, it shows a new panel, called QP, centered on the source at line 29
(Figure ~\ref{fig:example3}), the point of \textit{lastChange}.  To
the right, the TraceData panel shows values of properties of the
program state when it passed through line 29.  These two panels
resemble the Script and Watch panels, but they show data collected by
the debugger at one execution point which is now past: these are
'traces' or 'logs' of information collected during the re-execution.

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer reruns \textit{lastChange} command
on \texttt{myCondition.value} at this point. This re-executes the
program and pauses at the reproduction point,  showing line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value} (Figure ~\ref{fig:example4}). Now it is
clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once execution reaches line 25.

As demonstrated in Figure ~\ref{fig:example-points}, the developer has
examined three points of execution. The first point, the reproduction point,  was the breakpoint set by the developer.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure ~\ref{fig:example4}
there is an opened list which shows all three examined points. The
first one is the breakpoint on line 22, the second one is the point
which is when \texttt{myObject.myProperty} changed before
reaching the breakpoint and finally the last one is the point of
execution in which \texttt{myCondition.value} gets the
\texttt{undefined} value. Moreover, the source lines related to these
points are tagged with red-cycle icons.

@@

\subfigure[A screen shot of the Firebug debugger while running the example code from Fig.~\ref{fig:example2}. The Script
  panel is selected; it gives access to
  all loaded source files and allows breakpoints to be set on lines. In this
  figure, the execution is paused at line 22 by a regular
  breakpoint. The Watch panel at the right shows the program state at
  the paused
  point. ]{\label{fig:example1}\includegraphics[width=1.0\textwidth,
    height=.17\textheight]{1-bp22.jpg}}

\subfigure[Developer can query \textit{lastChange} on a value by right-clicking 
on the
  value of \texttt{myProperty}.]{\label{fig:example2}\includegraphics[width=1.0\textwidth,
    height=.17\textheight]{2-bp22-lastChange.jpg}}

\subfigure[The result of \textit{lastChange} query for
  \texttt{myObject.myProperty}. The left panel, QP, shows the source
  code at the point of \textit{lastChange}; The right panel,
  TraceData, shows the collected data at the
  point.]{\label{fig:example3}\includegraphics[width=1.0\textwidth,
    height=.21\textheight]{3-lastChange.jpg}}

\subfigure[The result of \textit{lastChange} query for
  \texttt{myCondition.value}. The opened list on the top of the left
  panel shows the visited execution points. Clicking on each point in
  the list shows the corresponding code and
  data.]{\label{fig:example4}\includegraphics[width=1.0\textwidth,
    height=.23\textheight]{4-lastChange2.jpg}}

@@

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[tight,normalsize]{subfigure}
%\usepackage{caption}
\usepackage{algorithmic}
\usepackage{algorithm}

@@

By browsing through the code, the developer can quickly determine that the value displayed on the 
button is set at line 22. To start debugging, the developer sets a breakpoint
on line 22. Once the button is clicked, execution is paused at line
22. Figure~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

@@

In contrast, we have added a high-level function in debugger,
\textit{lastChange}, which provides the answer without tedious manual
effort from the developer. By right clicking on
\texttt{myObject.myProperty} in the Watch panel, the developer can run
\textit{lastChange} command (Figure ~\ref{fig:example1}). Debugger
re-executes the program and halts again at the breakpoint on line 22.
However, it shows a new panel, called QP, centered on the source at line 29
(Figure~\ref{fig:example3}), the point of \textit{lastChange}.  To
the right, the TraceData panel shows values of properties of the
program state when it passed through line 29.  These two panels
resemble the Script and Watch panels, but they show data collected by
the debugger at one execution point which is now past: these are
'traces' or 'logs' of information collected during the re-execution.

Looking at line 29, it seems that something is wrong with
\texttt{myCondition.value} which causes line 29 execution. The
developer examines \texttt{myCondition.value} and it is
\texttt{undefined}. The next step is to know when this property got
this value. To do so, the developer reruns \textit{lastChange} command
on \texttt{myCondition.value} at this point. This re-executes the
program and pauses at the reproduction point,  showing line 25-the
place \texttt{oldValue} is assigned to
\texttt{myCondition.value}. To examine \texttt{oldValue}
 at this point, developer enters it to the watch box and after re-execution the value
 is available. Now it is clear that the bug occurs because \texttt{oldValue} is
\texttt{undefined} once execution reaches line 25 (Figure~\ref{fig:example4}).

@@

\subfigure[A screen shot of the Firebug debugger while running the example code from Fig.~\ref{fig:js-code}. The Script
  panel is selected; it gives access to
  all loaded source files and allows breakpoints to be set on lines. In this
  figure, the execution is paused at line 22 by a regular
  breakpoint. The Watch panel at the right shows the program state at
  the paused
  point. Developer can query \textit{lastChange} on a \texttt{myObject.myProperty} by right-clicking on the value of \texttt{myProperty}. ]{\label{fig:example1}\includegraphics[width=1.0\textwidth, height=.17\textheight
  ]{1-bp22.jpg}}

%\subfigure[Developer can query \textit{lastChange} on a value by right-clicking 
%on the
%  value of \texttt{myProperty}.]{\label{fig:example2}\includegraphics[width=1.0\textwidth,
%    height=.17\textheight]{2-bp22-lastChange.jpg}}

\subfigure[The result of \textit{lastChange} query for
  \texttt{myObject.myProperty}. The left panel, QP, shows the source
  code at the point of \textit{lastChange}; The right panel,
  TraceData, shows the collected data at the
  point.]{\label{fig:example3}\includegraphics[width=1.0\textwidth,% height=.21\textheight
  ]{3-lastChange.jpg}}

\subfigure[The result of \textit{lastChange} query for
  \texttt{myCondition.value}. To evaluate an expression (e.g., oldValue) at this point, developer can
  enter the expression in the watch box and after re-execution the result is available.
  The opened list on the top of the left panel shows the visited execution points. Clicking on each point in
  the list shows the corresponding code and
  data.]{\label{fig:example4}\includegraphics[width=1.0\textwidth%,height=.25\textheight
    ]{4-lastChange2.jpg}}

@@

As demonstrated in Figure ~\ref{fig:example-points}, the developer has
examined three points of execution. The first point, the reproduction point,  was the breakpoint set by the developer.
The second and third points preceded the reproduction point in execution sequence.
All three points-the history
of the search for the defect-are available through the debugger's
interface. On the top of the left panel in Figure~\ref{fig:example4}
there is an opened list which shows all three examined points. The
first one is the breakpoint on line 22, the second one is the point
which is when \texttt{myObject.myProperty} changed before
reaching the breakpoint and finally the last one is the point of
execution in which \texttt{myCondition.value} gets the
\texttt{undefined} value. Moreover, the source lines related to these
points are marked with red-cycle icons.
